---
title: "NYPD Shooting Information"
author: "Divya Leela Mohan"
date: "11/26/2021"
output:
  pdf_document: default
  html_document: default
---
##### Introduction: This assignment will show the ability to complete all steps in the data science process in a reproducible manner. I have to produce a report on the NYPD Shooting Incident data that includes your own analysis of some type.
### Step-1
set our working directory
```{r}
getwd()
```

```{r}
setwd("~/Rstudio Universty works/NYPD shooting Project")
```
install.packages("readr")
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(tidyverse)
```
##1.Importing Data

shooting<-read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")

NYPD_Shooting_Incident_Data_Historic_<-read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")
View(NYPD_Shooting_Incident_Data_Historic_)
```{r}
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
shooting <- read_csv(url)
```
##Cleaning
```{r}
shootinginfo = shooting[,-5, -6]
head(shootinginfo)
shootinginfo
summary(shootinginfo)
```

### Step 2-Tdying and Transforming Data
### method to change date  chr to date time

install.packages("chron")
install.packages("date")
```{r}
library(chron)
library(lubridate)
library(date)
```

class
OCCUR_DATE <- "08/27/2006" 
OCCUR_DATE data type has to transform from chr to datetype
```{r}
OCCUR_DATE <- as.Date("OCCUR_DATE","%m/%d/%y" )
class(myd(OCCUR_DATE))
head(shootinginfo)
summary(shootinginfo)
shootinginfo$OCCUR_DATE <- mdy(shootinginfo$OCCUR_DATE)
head(shootinginfo)
```


Occur_Date is changed to date format from the character format

## Step-3
## Analysis and visualization -1
### Q- Which boro has the highest number shooting incidents?

```{r}
shooting <- read_csv(url)
Shooting_boro <- shooting
Shooting_boro_count <- Shooting_boro %>% group_by(BORO) %>% count(BORO)
Shooting_boro_count
```
### Ans- From the analysis report its the Brooklyn which has the highest shooting incidents

### Analysis and visualization -2
### Q-Which race became the highest victim of shooting?

```{r}
shooting <- read_csv(url)
Shooting_vic_count <- shooting %>% group_by(VIC_RACE) %>% count(VIC_RACE)
Shooting_vic_count
summary(shootinginfo)
```

### Ans- From the analysis its the black who became the highest number of shooting victim


## visualization is visual representation of data in the form of graph or chart or bar or any other format-relation ship between data and images.
```{r}
library(tidyverse)
```

```{r}
ggplot(data = Shooting_boro_count,
    mapping =aes(x= n, y=BORO, group = 1 ))+
    geom_point(size= 3)+
    geom_line(color = "red")
```
   
    
### Another Sample for visualization plot
    
```{r}
ggplot(Shooting_boro_count,aes(BORO,n, group=1))+ geom_point()+geom_line()
```
### Visualization 3

```{r}
ggplot(Shooting_vic_count,aes( VIC_RACE, n, group=1))+ geom_point()+geom_line()+theme(text=element_text(size=8), axis.text.x = element_text(angle = 90))
```
#### Analysis-Black is the highest victim

## visualization 4
view(shooting) 
shooting
shootinginfo <- read_csv(url)
```{r}
shootinginfo %>% ggplot(aes(VIC_RACE,VIC_SEX,color= BORO))+
  geom_point(size=3, alpha=0.5)+
  theme(text=element_text(size=8), axis.text.x = element_text(angle = 90))+
  labs(x= "VIC_RACE", y= "VIC_SEX", title ="NYPD shooting")
```

### Step 4-Modelling 1.
```{r}
shooting <- read_csv(url)
view(shooting)
shooting %>% filter(VIC_AGE_GROUP< 50) %>%
  ggplot(aes(BORO,VIC_AGE_GROUP))+
  geom_point(aes(color = VIC_RACE,
  size= VIC_SEX),alpha =0.5)+ geom_smooth(method =lm)+
facet_wrap(~STATISTICAL_MURDER_FLAG,nrow=1)+
theme(text=element_text(size=8), axis.text.x = element_text(angle = 90))+
labs(x= "BORO", y= "Victim Age", title ="NYPD shooting")

```

### Conclusion and Identifying Bias
 
Newyork was some kind of a dream city and I never expect this much shootings from a borough.
before doing the analysis I thought all the races are victims equally. that was my bias, after doing the visualization and modelling, i came to realize that black race is the highest victim through out these years
If I go with my bias I couldnt get this result. in Vic_race.
Racial bias shooting is present from the given NYPD shooting historic report.
Statistical Murder flag True and false are separated so that reality is revealed

### Modelling 2- Incident_key and Precinct

```{r}
head(shooting)
sapply(shooting, class)
shooting$INCIDENT_KEY <-as.numeric(shooting$INCIDENT_KEY)
shooting$PRECINCT <- as.numeric(shooting$PRECINCT)
class("INCIDENT_KEY, PRECINCT")
plot(shooting$PRECINCT, shooting$INCIDENT_KEY, main= "Scatterplot")
cor(shooting$PRECINCT, shooting$INCIDENT_KEY)
help(lm)
?lm
#to finf mod
mod <- lm(shooting$INCIDENT_KEY ~ shooting$PRECINCT)
summary(mod)
attributes(mod)
mod$coefficients
coef(mod)
plot(shooting$PRECINCT, shooting$INCIDENT_KEY, main= "Scatterplot")
abline(mod)
abline(mod, col =2, lwd = 3)
plot(shooting$PRECINCT, shooting$INCIDENT_KEY, main= "Scatterplot")
class("INCIDENT_KEY, PRECINCT")
plot(shooting$PRECINCT, shooting$INCIDENT_KEY, main= "Scatterplot")
cor(shooting$PRECINCT, shooting$INCIDENT_KEY)
abline(mod, col =2, lwd = 3)
coef(mod)
confint(mod)
confint(mod, level =0.99)
summary(mod)
anova(mod)
sqrt(3.69)
```

### Identifying Bias
## Conclusion and Identifying Bias
 
Newyork was some kind of a dream city and I never expect this much shootings from a borough.
before doing the analysis I thought all the races are victims equally. that was my bias, after doing the visualization and modelling, i came to realize that black race is the highest victim through out these years
If I go with my bias I couldn't get this result. in Vic_race.
Racial bias shooting is present from the given NYPD shooting historic report.
 
##### My bias was in Vic_race. After prediction from the modelling that bias was identified and cleared.
